---
title: "River Crossing Puzzles"
author: "Mark Borg"
date: "`r Sys.setlocale('LC_ALL', 'English'); format(Sys.time(), '%d %B %Y')`"
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
getwd()
knitr::opts_knit$set(root.dir = getwd())

```


This R notebook is about **River Crossing Puzzles**.

For a description of these puzzles and how we go about creating the solution presented here, please see [this related blog entry](https://mark-borg.github.io/blog/2016/river-crossing-puzzles/).


```{r, message=FALSE, warning=FALSE}

source('solve_river_crossing_puzzles.R')

```


**Farmer-Wolf-Goat-Cabbage puzzle**

Setting up the incompatibilities between the different entities as a directed graph. E.g. Wolf (W) eats Goat (G).
And creating the search space illustrating all possible transitions (moves) that can be done.


```{r}

require(igraph)

# the graph showing object incompatibilities
gr <- make_empty_graph(directed = TRUE)
gr <- add.vertices(gr, 4, name = c('F', 'W', 'G', 'C'))
gr <- add.edges(gr, c('W','G', 'G','C'))
plot(gr)

# create the search space 
gss <- make_empty_graph(directed=TRUE)

boat.capacity <-2
farmer.symbol <- 'F'

# create the initial state
state0 <- list(bank.l = c('F', 'W', 'G', 'C'), bank.r = c(), boat.pos = 1)
state0 <- make.state.name(state0)

# add the initial state as a node in the search space
gss <- add.vertices(gss, 1, name=state0$name)
V(gss)[1]$color <- 'red'

gss <- solve(gss, state0)

gss <- simplify(gss, remove.loops = FALSE, remove.multiple = TRUE)

V(gss)[startsWith(V(gss)$name, '|')]$color <- 'green'
plot(gss)

```


Solving the puzzle by finding valid paths in the search space.


```{r}

paths <- get.all.shortest.paths(gss, 'CFGWb|', '|CFGWb')
print(paths)

paths <- get.all.shortest.paths(gss, V(gss)[endsWith(V(gss)$name, '|')], V(gss)[startsWith(V(gss)$name, '|')])
print(paths)

```


**Farmer-Fox-Chicken-Spider-Caterpillar-Lettuce**

Setting up the incompatibilities between the different entities as a directed graph. E.g. Fox (F) eats Chicken (C).
And creating the search space illustrating all possible transitions (moves) that can be done.


```{r}

# the graph showing object incompatibilities
gr <- make_empty_graph(directed = TRUE)
gr <- add.vertices(gr, 6, name = c('M', 'F', 'C', 'S', 'K', 'L'))
gr <- add.edges(gr, c('F','C', 'C','S', 'S', 'K', 'K', 'L'))
plot(gr)

# create the search space 
gss <- make_empty_graph(directed=FALSE)


boat.capacity <- 3
farmer.symbol <- 'M'

# create the initial state
state0 <- list(bank.l = c('M', 'F', 'C', 'S', 'K', 'L'), bank.r = c(), boat.pos = 1)
state0 <- make.state.name(state0)

# add the initial state as a node in the search space
gss <- add.vertices(gss, 1, name=state0$name)
V(gss)[1]$color <- 'red'

gss <- solve(gss, state0)

gss <- simplify(gss, remove.loops = FALSE, remove.multiple = TRUE)

V(gss)[startsWith(V(gss)$name, '|')]$color <- 'green'
plot(gss)

```


Solving the puzzle by finding valid paths in the search space.


```{r}

paths <- get.all.shortest.paths(gss, 'CFKLMSb|', '|CFKLMSb')
print(paths)

paths <- get.all.shortest.paths(gss, V(gss)[endsWith(V(gss)$name, '|')], V(gss)[startsWith(V(gss)$name, '|')])
print(paths)

```


**Two Wolves, a Dog, a Goat and a Bag of Grain**

Setting up the incompatibilities between the different entities as a directed graph. 
And creating the search space illustrating all possible transitions (moves) that can be done.


```{r}

# the graph showing object incompatibilities
gr <- make_empty_graph(directed = TRUE)
gr <- add.vertices(gr, 5, name = c('F', 'W', 'D', 'G', 'B'))
gr <- add.edges(gr, c('W','D', 'W','G', 'D', 'G', 'G', 'B'))
plot(gr)

# create the search space 
gss <- make_empty_graph(directed=FALSE)


boat.capacity <- 3
farmer.symbol <- 'F'

# create the initial state
state0 <- list(bank.l = c('F', 'W', 'W', 'D', 'G', 'B'), bank.r = c(), boat.pos = 1)
state0 <- make.state.name(state0)

# add the initial state as a node in the search space
gss <- add.vertices(gss, 1, name=state0$name)
V(gss)[1]$color <- 'red'

gss <- solve(gss, state0)

gss <- simplify(gss, remove.loops = FALSE, remove.multiple = TRUE)

V(gss)[startsWith(V(gss)$name, '|')]$color <- 'green'
plot(gss)

```


Solving the puzzle by finding valid paths in the search space.


```{r}

paths <- get.all.shortest.paths(gss, 'BDFGWWb|', '|BDFGWWb')
print(paths)

paths <- get.all.shortest.paths(gss, V(gss)[endsWith(V(gss)$name, '|')], V(gss)[startsWith(V(gss)$name, '|')])
print(paths)

```


**Missionaries and Cannibals**

For this problem, we need to consider counts of objects rather than incompatibilities between object types.
Thus we will ovveride some of the functions in `solve_river_crossing_puzzles.R` that we have used out-of-the-box for the previous puzzles.


```{r}

# for this problem, we need to consider counts of objects rather than incompatibilities between object types; thus override this method
is.bank.valid <- function(gr, state, side)
{
     b <- state[[side]]
     t <- table(b)
     num.c <- ifelse(is.na(t['C']), 0, t['C'])
     num.m <- ifelse(is.na(t['M']), 0, t['M'])
     return(num.m >= num.c | num.m == 0)
}

# for this problem, the only rule is that the boat is not empty; thus override this method
is.transition.valid <- function(transition)
{
     return(min(is.na(transition)) == 0)
}

```


We create the search space illustrating all possible transitions (moves) that can be done.


```{r}

# the graph showing object incompatibilities; not used in this particular case - we will leave it empty
gr <- make_empty_graph(directed = TRUE)

# create the search space 
gss <- make_empty_graph(directed = FALSE)


boat.capacity <- 2

# create the initial state
state0 <- list(bank.l = c('M', 'M', 'M', 'C', 'C', 'C'), bank.r = vector(), boat.pos = 1)
state0 <- make.state.name(state0)

# add the initial state as a node in the search space
gss <- add.vertices(gss, 1, name=state0$name)
V(gss)[1]$color <- 'red'

gss <- solve(gss, state0)

gss <- simplify(gss, remove.loops = FALSE, remove.multiple = TRUE)

V(gss)[startsWith(V(gss)$name, '|')]$color <- 'green'
plot(gss)

```


Solving the puzzle by finding valid paths in the search space.


```{r}

paths <- get.all.shortest.paths(gss, 'CCCMMMb|', '|CCCMMMb')
print(paths)

```


**Japanese Family River Crossing Puzzle**

Again, for this puzzle we need to override some of the standard methods in file `solve_river_crossing_puzzles.R`.


```{r}

# for this problem, we need to consider complex incompatibilities between object types; thus override this method
is.bank.valid <- function(gr, state, side)
{
    b <- state[[side]]
    if (! is.element('M',b) & length(b) > 1 & is.element('F',b) & is.element('D',b))     { return (FALSE) }
    if (! is.element('F',b) & length(b) > 1 & is.element('M',b) & is.element('S',b))     { return (FALSE) }
    if (! is.element('P',b) & length(b) > 1 & is.element('T',b) & (is.element('F',b) | is.element('M',b) | is.element('S',b) | is.element('D',b)))     { return (FALSE) }

    return(TRUE)
}

# for this problem, multiple persons can operate the boat; thus override this method
is.transition.valid <- function(transition)
{
     return(is.element('M', transition) | is.element('F', transition) | is.element('P', transition))
}

```

We create the search space illustrating all possible transitions (moves) that can be done.


```{r}

# the graph showing object incompatibilities; not used in this particular case
gr <- make_empty_graph(directed = TRUE)

# create the search space 
gss <- make_empty_graph(directed = FALSE)


boat.capacity <- 2

# create the initial state
state0 <- list(bank.l = c('F', 'M', 'P', 'T', 'D', 'D', 'S', 'S'), bank.r = vector(), boat.pos = 1)
state0 <- make.state.name(state0)

# add the initial state as a node in the search space
gss <- add.vertices(gss, 1, name=state0$name)
V(gss)[1]$color <- 'red'

gss <- solve(gss, state0)

gss <- simplify(gss, remove.loops = FALSE, remove.multiple = TRUE)

V(gss)[startsWith(V(gss)$name, '|')]$color <- 'green'
plot(gss, layout=layout.fruchterman.reingold(gss, niter=10000), vertex.label.cex=0.6)

```


Solving the puzzle by finding valid paths in the search space.


```{r}

paths <- get.all.shortest.paths(gss, 'DDFMPSSTb|', '|DDFMPSSTb')

```


